Projection variation: Is the family of sentences really a family?

### 2024-02-21 writing/SuB-paper/projection-operators-SuB.tex
tracking changes from overleaf

\usepackage{subcaption}

@@

\usepackage{titleps}
\usepackage{lipsum}
\newpagestyle{mystyle}{
\sethead[\thepage][\Author][]{}{\Title}{\thepage}
}
\pagestyle{mystyle}
\author{Hofmann \and de Marneffe  \and Tonhauser}
%\title{Projection variability of clausal complements across different operators}
\title{Projection variability: Is the family of sentences really a family?}
\makeatletter
\newcommand\Author{Hofmann, de Marneffe, and Tonhauser}
\let\Title\@title
\makeatother
%header with author and title

@@

% drawing
	\usepackage{graphicx}
	\usepackage{xcolor}
    \definecolor{op-n-color}{HTML}{56B4E9}
    \definecolor{op-m-color}{HTML}{E69F00}
    \definecolor{op-c-color}{HTML}{009E73}
    \definecolor{op-q-color}{HTML}{0072B2}
    \definecolor{pred-fcv-color}{HTML}{CC79A7}

@@

{\large \textbf{Projection variability: Is the family of sentences really a family?}}\footnote{We thank Taylor Mahler for assistance in collecting the data presented here as well as valuable comments. We gratefully acknowledge the National Science foundation grant \#1452674 (to Marie-Catherine de Marneffe, Craige Roberts, and Judith Tonhauser), which provided financial support for this project.}\\
Lisa HOFMANN --- \textit{University of Stuttgart}\\
Marie-Catherine DE MARNEFFE --- \textit{UCLouvain}\\
Judith TONHAUSER --- \textit{University of Stuttgart}\\

\textbf{Abstract.} We present experimental evidence \textbf{(i)} that the projection of the content of clausal complements  varies across entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from non-factives predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents. The observed variability is not captured by existing theoretical accounts of projection (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}). Our results suggest that an analysis must consider interactions between predicates and operators and raise important questions for future research on projection.

\textbf{Keywords:} Projection variability, entailment-canceling operators, (non)factive predicates.

\section{Introduction} \label{sec:intro}
    The so-called `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective in formal semantics/pragmatics (e.g., \citealt{chierchia_meaning_1990}). For instance, the clausal complement of \emph{discover} is diagnosed as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from the variants in \ref{ex:family}, where \ref{ex:simple} is embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, an epistemic possibility modal \ref{ex:mod}, or in a conditional antecedent \ref{ex:cond}.

    % In formal semantics/pragmatics, the so-called `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective (e.g., \citealt{chierchia_meaning_1990}). For instance, to diagnose the content of the clausal complement of \emph{discover} as projective content, a researcher would typically observe that the content that Julian dances salsa follows not just from \ref{ex:simple}, but also from the family-of-sentences variants of \ref{ex:simple} in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.

@@

    Some research, however, suggests that entailment-cancelling operators may affect projection differentially. For instance, \citet{karttunen_observations_1971} proposed distinguishing English factive predicates (e.g., \textit{regret}) from semi-factives (e.g., \textit{discover}). Based on \Next, he argued that the content of the complement (CC) of true factives consistently projects across the operators in \Last, but that of semi-factives does not always project from under polar questions, modals, or conditionals.%

@@

    There have been two experimental investigations of by-operator projection variation. First, \citet{smith_relationship_2014} investigated projection from under negation and conditional antecedents for various types of English projective contents. They found that the expressive content of epithets (e.g., \textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals. In contrast, the content of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and the existential presupposition of clefts showed no difference.
    %However, their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance. Since other factors besides speaker commitment modulate discourse (un)expectedness (see e.g., \citealt{zimmermann_grammatical_2011,tonnis_german_2021}), it is unclear that they (only) measured projection.

    Second, \citet{sieker_projective_2022}  compared the projection of the CCs of German factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{ent\-hüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') from under the four operators in \ref{ex:family}. The results suggest that the CCs project more from under negation than from under the other three operators. Contrary to what \citet{karttunen_observations_1971} suggested, a comparison of the factive and semi-factive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of semi-factive predicates.\footnote{The factive/semi-factive distinction is also called into question by naturally occurring examples where the CCs of factive predicates do not project from under the four operators (see \citealt{beaver_have_2010,de_marneffe_did_2012,de_marneffe_commitmentbank_2019}). For experimental research on the distinction, see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}.
    }

    %They replicated \poscite{smith_relationship_2014} result that the CC of \textit{know} projects more from under negation than conditionals for German \textit{wissen} (`know'). However, this finding was part of an overall pattern of higher projection ratings with negation than other operators.
    %
    %Grouping ratings for the four factive and semi-factive predicates, respectively, they found no interaction of operator with predicate type, and thus no evidence for the factive/semi-factive distinction (iii), or an operator/predicate interaction (ii). However, their analysis did not explore the impact of operators on each predicate individually.

@@

    The results of our study suggest that the projection of the CCs of these 20 predicates is affected differently by the four entailment-cancelling operators in \ref{ex:family},  but not in a way that is consistent with \poscite{karttunen_observations_1971} factive/semi-factive distinction. The results also replicate a result of \citealt{degen_are_2022}, namely that projection from under polar questions does not categorically distinguish factive and non-factive predicates. Finally, the results of our experiment extend this result of \citealt{degen_are_2022} to projection from under the other three entailment-canceling operators, thereby solidifying their claim that ``research on projective content has a much broader empirical scope than previously assumed'' (p.585), as this scope includes the CCs of both factive and non-factive predicates.

\section{Experiments}\label{s2}
	To assess the effect of entailment-cancelling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.
    %
    The predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal {\em perhaps} in Exps.~3, and in conditional antecedents in Exps.~4. Each set of experiments contained three experiments, that differed in an at-issueness measure used in a separate block. In this paper, we limit our attention to the projection ratings collected in these twelve experiments.\footnote{\label{footnote-supplement}The experiments, data and analysis scripts, as well as supplements referred to in this paper can be found in the following GitHub repository: \url{https://github.com/judith-tonhauser/CommitmentBankPlus}.}

	In all twelve experiments, projection was measured with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019,sieker_projective_2022}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.}
    %This diagnostic was also used in \citealt{sieker_projective_2022}.
    Under this diagnostic, participants are presented with utterances like those in \Next, and asked to rate whether the (named) speaker is certain of the CC.

@@

    We assume, following \citet{tonhauser_how_2018} and \citet{degen_are_2022}, that judgments of speaker certainty about the embedded content reflect speaker commitment to that content, that is, projection. If a participant interprets utterances like \Last[a--d] in a way that the speaker (here, Christopher) is certain of the CC, the CC is assumed to project. If a participant does not take the speaker to be certain of the CC, the CC is taken to not project.

    %This setup allows us to test the three hypotheses outlined in Section \ref{sec:intro}. First, a main effect of operator on speaker certainty judgments for embedded contents would provide evidence for by-operator variation (i). Second, if there is a operator/predicate interaction (ii), we would expect the effect of operator to be different (in magnitude or direction) for different predicates. Finally, based on the factive/semi-factive distinction (iii), we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.

	\subsection{Methods}

		%\subsubsection{Participants}
        \paragraph{Participants.}
			We recruited 250-300 participants for each of the 12 experiments. Participants for one experiment were recruited on Amazon's Mechanical Turk platform. These participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs. Participants for the remaining experiments were recruited on Prolific. These participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%. See Supplement \ref{app:d-participants} (in the repository linked to in footnote \ref{footnote-supplement}) for further information about the participants.

            \begin{figure}[h!]
				\centering
				\fbox{\includegraphics[width = .6\linewidth]{task-1n-proj.eps}}
				\caption{A sample trial in Exps.~1. In the corresponding trials in the other experiments, participants were presented with an utterance with a different entailment-cancelling operator.}
				\label{fig:trial}
			\end{figure}

		% \subsubsection{Materials}
        \paragraph{Materials.}
            The target sentences consisted of the 400 combinations of the 20 clause-em\-bed\-ding predicates in \ref{ex:predicates} with 20 embedded clauses (provided in Supplement \ref{app:a-clauses}). As mentioned above, the predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal \emph{perhaps} in Exps.~3, and in conditional antecedents in Exps.~4, for a total of 400 target stimuli in each of the four sets of experiments. To assess whether participants were attending to the task, each experiment included six control stimuli. For details on the six control stimuli, see Supplement \ref{app:c-control}.

            Each participant saw a random set of twenty-six stimuli: Each set contained one target stimulus for each of the 20 clause-embedding predicates (each with a unique complement clause) and the same six control stimuli.\footnote{Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block. Block order was randomized. As mentioned above, we focus on the projection ratings here.} Trial order was randomized.

        %\subsubsection{Procedure}
        \paragraph{Procedure.}
            Participants were asked to imagine that they are at a party and that, when walking into the kitchen, they overhear somebody say something to somebody else.
            %
            On each trial, they read an utterance and gave a response to the `certain that' question on a slider marked `no' (coded as 0) at one end and `yes' (coded as 1) on the other. A sample trial from Exps.~1 is shown in Figure~\ref{fig:trial}.

            Following \citet{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things: On the one hand, higher ratings could reflect greater speaker commitment towards the CC, and therefore greater projection. This assumes that speaker commitment is interpreted in a gradient way.
            % gradient.
            Conversely, higher certainty ratings could reflect the probability that an interpreter takes the speaker to be committed to the CC.  On this interpretation, speaker commitment may be a binary, categorical property and projection variation is a result of the uncertainty about possible interpretations. In this paper, we remain agnostic about the underlying interpretation of projectivity as a gradient property, though our discussion of projection variability will be in line with the first interpretation (for discussion, see also \citealt{grove_factivity_2023}).

@@

        %\subsubsection{Data exclusion}
        \noindent\textbf{Data exclusion}
            was done based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}. The data from 2,682 participants entered into the analysis.
            % Data were excluded based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}. The data from 2,682 participants entered into the analysis.

    \subsection{Results and discussion}

        We first address the question of whether there is by-operator variation (section \ref{s:by-op}) and then the question of whether there is by-predicate variation in the observed by-operator variation (section \ref{s:by-pred}). Finally, in section \ref{s:factivity}, we relate the results to the result of \citealt{degen_are_2022}, that projection from under polar questions does not categorically distinguish factive and non-factive predicates.

		%Our data reveals three key results: \textbf{(i)} There is projection variability by operator; \textbf{(ii)} There is by-predicate variation in the effect of operator on projection; and \textbf{(iii)} We find further support (from entailment-cancelling operators beyond polar questions) for \posscite{degen_are_2022} result that by-predicate projection variation is gradient in a way that cannot be predicted by categorical lexical classes.

	\subsubsection{By-operator variation}\label{s:by-op}

        Figure~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates. As shown, there is projection variability by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relatively more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.

            \begin{figure}[ht]
				\centering
				\includegraphics[scale = .8]{certainty-operator.pdf}
				\caption{Mean certainty ratings by operator. Error bars indicate $95\%$ bootstrapped confidence intervals and violin plots indicate the kernel probability density of the ratings.
                % of the individual participants’ ratings.
                }
				\label{fig:op-ratings}
			\end{figure}

        These generalizations are supported by a post-hoc pairwise comparison of the estimated means for each entailment-canceling operator using the \texttt{emmeans} package (\citealt{emmeans}) in R (\citealt{r}). The input to the pairwise comparison was a  Bayesian mixed-effects beta regression model that was fit using the \texttt{brms} package (\citealt{buerkner2017}) with weakly informative priors. The model predicted certainty ratings\footnote{To model the certainty ratings using a beta regression, the ratings were first transformed from the interval [0,1] to the interval (0,1) using the method proposed in \citealt{smithson_better_2006}.} from a fixed effect of entailment-canceling operator (with treatment coding and `modal' as the reference level) and included a random by-predicate intercept.\footnote{See Supplement \ref{app:analysis-1} for more details on the model.} The output of the pairwise comparison were 95\% highest density intervals (HDIs) of estimated marginal mean differences between each of the operators. We assume that two operators differ in certainty ratings if the HDI of their pairwise comparison does not include 0.

        Table \ref{t:pairwise} provides the output of the pairwise comparison on a logit scale.
        %\footnote{While the certainty ratings collected from the participants range from 0 to 1, the smaller than 0 estimated marginal mean differences are the result of the beta regression using a logit link function for the mean parameter.}
        As shown, the analysis suggests differences between each pair of operators, though the estimated marginal mean differences differ across the pairs. That is, certainty ratings are higher for CCs embedded in conditional antecedents than those embedded under polar questions; certainty ratings for CCs embedded under polar questions are higher than those for CCs embedded under negation; and finally certainty ratings for CCs embedded under negation are higher than those for CCs embedded under the epistemic modal \emph{perhaps}.

        \begin{table}[!h]
        \centering
        \begin{tabular}{lrrr}
         contrast & estimate & lower 95\% CI & upper 95\% CI \\\hline 
         conditional - negation & 0.21 & 0.18 & 0.24 \\ 
          conditional - question & 0.13 & 0.10 & 0.16 \\ 
          modal - conditional & -0.26 & -0.29 & -0.23 \\ 
          modal - negation & -0.05 & -0.08 & -0.02 \\ 
          modal - question & -0.14 & -0.17 & -0.11 \\ 
          negation - question & -0.09 & -0.12 & -0.06 \\ 
        \end{tabular}
        \caption{Output of the pairwise comparison of entailment-canceling operators. The `contrast' column identifies the pair of entailment-canceling operators, the `estimate' column the estimated marginal mean difference, and the `lower 95\% CI' and `upper 95\% CI' columns provide the lower and upper bounds of the HDIs.}\label{t:pairwise}
        \end{table}

 These results suggest that certainty ratings for the CCs of of the English clause-embedding predicate we investigated vary by entailment-canceling operator. In contrast to \citet{sieker_projective_2022} for German, the results of our experiments do not suggest that projection is strongest from under negation. Recall, however, that they only investigated projection of the CCs of (semi-)factive predicates. Since there is by-predicate variation in the effect of entailment-canceling operator on projection (as we show in the next section), this difference between the results of their experiment and ours might very well be due to the types of predicates investigated. Finally, the differences in mean certainty ratings between the four entailment-canceling operators are very small. This suggests that, when abstracting away from individual predicates and contents, projection from under the four entailment-canceling operators is very similar. In other words, when abstracting away from individual  contents, the family of sentences really are a family.

	\subsubsection{By-predicate variation in the effect of entailment-canceling operator}\label{s:by-pred}

        Figure~\ref{fig:op-pred-analysis} shows mean certainty ratings by entailment-canceling operator for the 20  predicates, with predicates ordered by their overall mean certainty rating.
        %
        As shown, there is by-operator projection variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation. Further, the effect of operator differs between predicates. Examining the observed variation suggests that it does not align with widely assumed predicate classes.

         \begin{figure}[ht!]
            \centering
    		\includegraphics[width = .9\linewidth]{analysis.pdf}
    		\caption{Mean certainty ratings by predicate (\textcolor{pred-fcv-color}{(semi-)factive}, non-factive) and operator (\textcolor{op-m-color}{modal},  \textcolor{op-n-color}{negation}, \textcolor{op-q-color}{polar questions}, \textcolor{op-c-color}{conditional antecedents}) with 95\% bootstrapped confidence intervals. Violin plots indicate kernel probability density of the individual participants' ratings. Facets for predicates are ordered by the predicate mean certainty rating (aggregating across operators). Below each facet, a line spanning two operators indicates a non-zero difference according to the pairwise comparison of operators. The linetype indicates whether the difference is
            at least $1$ (solid line: —),
            below $1$ but at least $0.5$ (dashed line: – –),
            or below $0.5$ but above $0$ (dotted line: . . . )}
    		\label{fig:op-pred-analysis}
    	\end{figure}

>>>>

\usepackage{subcaption}
\usepackage{booktabs}

@@

\usepackage{titleps}
\usepackage{lipsum}
\newpagestyle{mystyle}{
\sethead[\thepage][\Author][]{}{\Title}{\thepage}
}
\pagestyle{mystyle}
\author{Hofmann \and de Marneffe  \and Tonhauser}
%\title{Projection variation of clausal complements across different operators}
\title{Projection variation: Is the family of sentences really a family?}
\makeatletter
\newcommand\Author{Hofmann, de Marneffe, and Tonhauser}
\let\Title\@title
\makeatother
%header with author and title

@@

% drawing
	\usepackage{graphicx}
	\usepackage{xcolor}
    \definecolor{op-n-color}{HTML}{009E73}
    \definecolor{op-m-color}{HTML}{0072B2}
    \definecolor{op-c-color}{HTML}{404040}
    \definecolor{op-q-color}{HTML}{E69F00}
    \definecolor{pred-fcv-color}{HTML}{CC79A7}

@@

{\large \textbf{Projection variation: Is the family of sentences really a family?}}\footnote{We thank Taylor Mahler for assistance in collecting the data presented here as well as valuable comments. We gratefully acknowledge the National Science Foundation grant BCS \#1452674 (to Marie-Catherine de Marneffe, Craige Roberts, and Judith Tonhauser), which provided financial support for this project.}\\
Lisa HOFMANN --- \textit{University of Stuttgart}\\
Marie-Catherine DE MARNEFFE --- \textit{UCLouvain}\\
Judith TONHAUSER --- \textit{University of Stuttgart}\\

\textbf{Abstract.}
    Under the `family of sentences' diagnostic for projection, the projection of content is investigated by embedding the expression that contributes the content in the scope of negation, polar questions, epistemic possibility modals, and conditional antecedents. This paper reports on the results of a set of experiments designed to investigate whether there is variation in the projection of content from under these four types of entailment-canceling operators. The contents investigated are the contents of the complements of 20 English clause-embedding predicates. The results of the experiments suggest (i) that the by-operator variation is small when aggregating over the 20 contents, but (ii) that the effect of operator differs between the clause-embedding predicates. The results of these experiments also extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from non-factive predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents. The observed by-predicate and by-operator variation is not captured by existing theoretical accounts of projection (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}). Our results suggest that an empirically adequate projection analysis must consider interactions between predicates and operators.

\textbf{Keywords:} Projection variation, entailment-canceling operators, (non)factive predicates.

\section{Introduction} \label{sec:intro}
    The `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective (e.g., \citealt{chierchia_meaning_1990}). For instance, in \ref{ex:simple}, the content of the clausal complement of \emph{discover} (that Julian dances salsa) is diagnosed as projective content, if it is typically implied not just by an utterance of \ref{ex:simple}, but also by utterances of the variants in \ref{ex:family}, where \ref{ex:simple} is embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, an epistemic possibility modal \ref{ex:mod}, or in a conditional antecedent \ref{ex:cond}.

@@

    Some research, however, suggests that entailment-canceling operators may affect projection differentially. For instance, \citet{karttunen_observations_1971} proposed distinguishing English factive predicates (e.g., \textit{regret}) from semi-factives (e.g., \textit{discover}). Based on \Next, he argued that the content of the complement (CC) of true factives consistently projects across the operators in \Last, but that of semi-factives does not always project from under polar questions, modals, or conditionals.%

@@

    There have been two experimental investigations of by-operator projection variation. First, \citet{smith_relationship_2014} investigated projection from under negation and conditional antecedents for various types of English projective contents. They found that the expressive content of epithets (e.g., \textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals. In contrast, the content of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and the existential presupposition of clefts showed no difference.

    Second, \citet{sieker_projective_2022}  compared the projection of the CCs of German factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{ent\-hüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') from under the four operators in \ref{ex:family}. Their results suggest that the CCs project more from under negation than from under the other three operators. Contrary to what \citet{karttunen_observations_1971} suggested, a comparison of the factive and semi-factive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of semi-factive predicates.\footnote{The factive/semi-factive distinction is also called into question by naturally occurring examples where the CCs of factive predicates do not project from under the four operators (see \citealt{beaver_have_2010,de_marneffe_did_2012,de_marneffe_commitmentbank_2019}). For experimental research on the distinction, see \citealt{djarv_cognitive_2018}.
    }

@@

    The results of our investigation suggest that the projection of the CCs of these 20 predicates is affected differently by the four entailment-canceling operators in \ref{ex:family},  but not in a way that is consistent with \poscite{karttunen_observations_1971} factive/semi-factive distinction. The results also replicate a result of \citealt{degen_are_2022}, namely that projection from under polar questions does not categorically distinguish factive and non-factive predicates. Finally, the results of our experiment extend this result of \citealt{degen_are_2022} to projection from under the other three entailment-canceling operators, thereby solidifying their claim that ``research on projective content has a much broader empirical scope than previously assumed'' (p.585), as this scope includes the CCs of both factive and non-factive predicates.

\section{Experiments}\label{s2}
	To assess the effect of entailment-canceling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.
    %
    The predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal {\em perhaps} in Exps.~3, and in conditional antecedents in Exps.~4. Each set of experiments contained three experiments that differed in the at-issueness measure that was used in a separate block. In this paper, we limit our attention to the projection ratings collected in these twelve experiments.\footnote{\label{footnote-supplement}The experiments, data and analysis scripts, as well as the supplements referred to in this paper can be found in the following GitHub repository: \url{https://github.com/judith-tonhauser/CommitmentBankPlus}.}

	In all twelve experiments, projection was measured with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019,sieker_projective_2022}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}, and discussion in \citealt{tonhauser_how_2018}.}
    Under this diagnostic, participants are presented with utterances like those in \Next, and asked to rate whether the (named) speaker is certain of the CC.

@@

    We assume, following \citealt{tonhauser_how_2018} and \citealt{degen_are_2022}, that judgments of speaker certainty about the embedded content reflect speaker commitment to that content, that is, projection. If a participant interprets utterances like \Last[a--d] in a way that the speaker (here, Christopher) is certain of the CC, the CC is assumed to project. If a participant does not take the speaker to be certain of the CC, the CC is taken to not project.

    \paragraph{Participants.}
	    We recruited 250-300 participants for each of the 12 experiments. Participants for one experiment were recruited on Amazon's Mechanical Turk platform. These participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs. Participants for the remaining experiments were recruited on Prolific. These participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%. See Supplement D (in the repository linked to in footnote \ref{footnote-supplement}) for further information about the participants.

            \begin{figure}[h!]
				\centering
                    %\fbox{\includegraphics[scale=0.5]{task-1n-proj.eps}}
				\fbox{\includegraphics[width = .6\linewidth]{task-1n-proj.eps}}
				\caption{A sample trial from Exps.~1 (negation). In the other experiments, participants were presented with utterances with a different entailment-canceling operator.}
				\label{fig:trial}
			\end{figure}

		% \subsubsection{Materials}
        \paragraph{Materials.}
            The target sentences consisted of the 400 combinations of the 20 clause-em\-bed\-ding predicates in \ref{ex:predicates} with 20 embedded clauses (provided in Supplement A). As mentioned above, the predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal \emph{perhaps} in Exps.~3, and in conditional antecedents in Exps.~4, for a total of 400 target stimuli in each of the four sets of experiments. To assess whether participants were attending to the task, each experiment included six control stimuli. For details on the six control stimuli, see Supplement C.

            Each participant saw a random set of 26 stimuli: Each set contained one target stimulus for each of the 20 clause-embedding predicates (each with a unique complement clause) and the same six control stimuli.\footnote{Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block. Block order was randomized. As mentioned above, we focus here on the projection ratings.} Trial order was randomized.

        %\subsubsection{Procedure}
        \paragraph{Procedure.}
            Participants were asked to imagine that they are at a party and that, when walking into the kitchen, they overhear somebody say something to somebody else.
            %
            On each trial, they read an utterance and gave a response to the `certain that' question on a slider marked `no' (coded as 0) at one end and `yes' (coded as 1) at the other. A sample trial %from Exps.~1
            is shown in Figure~\ref{fig:trial}.            
            Following \citealt{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things. First, higher certainty ratings could reflect greater speaker commitment towards the CC, and therefore greater projection. This assumes that speaker commitment is interpreted in a gradient way. Second, higher certainty ratings could reflect a higher probability that an interpreter takes the speaker to be committed to the CC.  On this interpretation, speaker commitment may be a binary, categorical property and projection variation is a result of uncertainty about speaker commitment. In this paper, we remain agnostic about the underlying interpretation of projectivity as a gradient property (for discussion, see  \citealt{grove_factivity_2023}).

@@

        %\subsubsection{Data exclusion}
        \noindent\textbf{Data exclusion.} Data were excluded based on self-declared non-native speaker status and other criteria given in Supplement D. The data from 2,682 participants entered into the analysis.
            % Data were excluded based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}. The data from 2,682 participants entered into the analysis.

\section{Results and discussion}

        We first address by-operator variation (Section \ref{s:by-op}) and then the question of whether there is by-predicate variation in the observed by-operator variation (Section \ref{s:by-pred}). Finally, in Section \ref{s:factivity}, we relate our results to those of \citealt{degen_are_2022}, that projection from under polar questions does not categorically distinguish factive and non-factive predicates.

	\subsection{By-operator variation}\label{s:by-op}

        Figure~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates. As shown, there is projection variation by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relatively more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.

            \begin{figure}[ht]
				\centering
				\includegraphics[scale = .7]{certainty-operator.pdf}
				\caption{Mean certainty ratings by operator. Error bars indicate $95\%$ bootstrapped confidence intervals. Violin plots indicate the kernel probability density of participants' individual ratings.
                }
				\label{fig:op-ratings}
			\end{figure}

        These observations are supported by a post-hoc pairwise comparison of the estimated means for each entailment-canceling operator using the \texttt{emmeans} package (\citealt{emmeans}) in R (\citealt{r}). The input to the pairwise comparison was a Bayesian mixed-effects beta regression model that was fit using the \texttt{brms} package (\citealt{buerkner2017}) with weakly informative priors. The model predicted certainty ratings\footnote{To model the certainty ratings using a beta regression, the ratings were first transformed from the interval [0,1] to the interval (0,1) using the method proposed in \citealt{smithson_better_2006}.} from a fixed effect of entailment-canceling operator (with treatment coding and `modal' as the reference level) and included a random by-predicate intercept.\footnote{See Supplement E for details on the model.} The output of the pairwise comparison were 95\% highest density intervals (HDIs) of estimated marginal mean differences between each of the operators. We assume that two operators differ in certainty ratings if the HDI of their pairwise comparison does not include 0.

        Table~\ref{t:pairwise} provides the output of the pairwise comparison on a logit scale.
        As shown, the analysis suggests differences between each pair of operators. That is, certainty ratings are higher for CCs embedded in conditional antecedents than those embedded under polar questions, certainty ratings for CCs embedded under polar questions are higher than those for CCs embedded under negation, and finally certainty ratings for CCs embedded under negation are higher than those for CCs embedded under the epistemic modal \emph{perhaps}.

        \begin{table}[!h]
        \centering
        \begin{tabular}{lrrr}
        \toprule
         contrast & estimate & lower 95\% CrI & upper 95\% CrI \\\midrule
         conditional - negation & 0.21 & 0.18 & 0.24 \\ 
          conditional - question & 0.13 & 0.10 & 0.16 \\ 
          modal - conditional & -0.26 & -0.29 & -0.23 \\ 
          modal - negation & -0.05 & -0.08 & -0.02 \\ 
          modal - question & -0.14 & -0.17 & -0.11 \\ 
          negation - question & -0.09 & -0.12 & -0.06 \\ \bottomrule
        \end{tabular}
        \caption{Output of the pairwise comparison of entailment-canceling operators. The `contrast' column identifies entailment-canceling operators pairs, `estimate' the estimated marginal mean difference, and `lower/upper 95\% CrI' provide the lower/upper bounds of the HDIs.}\label{t:pairwise}
        \end{table}

        These results suggest that certainty ratings for the CCs of the English clause-embedding predicates we investigated vary by entailment-canceling operator. In contrast to \citealt{sieker_projective_2022} for German, the results of our experiments do not suggest that projection is strongest from under negation. Recall, however, that they only investigated projection of the CCs of (semi-) factive predicates. Since there is by-predicate variation in the effect of entailment-canceling operator on projection (as we show in the next section), this difference between the results of their experiment and ours might be due to the types of predicates investigated. Finally, the differences in mean certainty ratings between the four entailment-canceling operators are very small. This suggests that, when abstracting away from individual predicates and contents, projection from under the four entailment-canceling operators is very similar. In other words, when abstracting away from individual  contents, the family of sentences really are a family.

	\subsection{By-predicate variation in the effect of entailment-canceling operator}\label{s:by-pred}

        Figure~\ref{fig:op-pred-analysis} shows mean certainty ratings by entailment-canceling operator for the 20 predicates, with predicates ordered by their overall mean certainty rating. As shown, there is by-operator projection variation for each predicate. Further, the effect of operator differs between predicates. For instance, the five (semi-)factive predicates exhibit four different patterns.
        %
        First, the CC of \emph{be annoyed} projects
        most from under questions, less from under negation, followed by conditionals, and least from under the modal \emph{perhaps} (\textcolor{op-q-color}{Q} $>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-c-color}{C} $>$ \textcolor{op-m-color}{M}).
        % more from under negation and questions than conditionals and the modal {\em perhaps};
        Second, the CC of \emph{know} projects most from under questions, less from conditionals and negation, and least from under {\em perhaps} (\textcolor{op-q-color}{Q} $> \{$\textcolor{op-n-color}{N}$,$ \textcolor{op-c-color}{C}$\} >$ \textcolor{op-m-color}{M}).
        %
        The CCs of \emph{discover} and \emph{see} exhibit a third pattern: They project most from under questions and conditionals, less from under negation, and least from under \emph{perhaps} ($\{$\textcolor{op-q-color}{Q}$,$ \textcolor{op-c-color}{C}$\}>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-m-color}{M}).
        %
        Finally, the CC of \emph{reveal} projects most from conditionals, less from questions, and least from under negation and \emph{perhaps} (\textcolor{op-c-color}{C} $>$ \textcolor{op-q-color}{Q}  $> \{$\textcolor{op-m-color}{M}$, $\textcolor{op-n-color}{N}$\}$).
        %
        Thus, the CCs of none of our (semi-)factive predicts project uniformly from under all four entailment-canceling operators (contrary to what \citealt{karttunen_observations_1971} suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators.
